---
title: "FTDR Assignment 1"
author: "Jesse Nieuwkoop (1689959)"
date: "28-11-2024-"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      expand: 3
    df_print: paged
---

---

# Preparation

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
# Import the libraries
library(tidyverse)    # For data manipulation
library(ggplot2)      # For plotting
library(ggpubr)       # For ggplot grid arrange
library(kableExtra)   # For fancy tables
```

## Read data

```{r}
# Read data
data <- read.csv("Airbnb_Open_Data.csv")

# View data structure
dplyr::glimpse(data)
```

# Preprocessing

## Duplicates

```{r}
# Print total duplicates
sum(duplicated(data$host.id))
```

```{r}
# Remove duplicate rows 
data <- data %>% 
  dplyr::distinct(host.id, .keep_all = TRUE)
```

## Subset & Rename

```{r}
# Subset data
data <- data %>%
  dplyr::select(cancellation_policy,
                room.type,
                Construction.year,
                service.fee,
                minimum.nights,
                number.of.reviews,
                reviews.per.month,
                review.rate.number,
                availability.365,
                price)

# Rename column names to snake_case format
data <- data %>%
  dplyr::rename(
    room_type = room.type,
    construction_year = Construction.year,
    service_fee = service.fee,
    minimum_nights = minimum.nights,
    number_of_reviews = number.of.reviews,
    reviews_per_month = reviews.per.month,
    review_rate_number = review.rate.number,
    availability_365 = availability.365,
    price_usd = price
  )
```

## Missing values

```{r}
# Count missing data
data %>%
  dplyr::summarise(across(everything(), ~ sum(is.na(.)))) %>%
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "NA Count"
  )
```

```{r}
# Drop all rows that contain empty or missing values
data <- data %>%
  # Convert empty character variables to NA
  mutate(across(where(is.character), ~ trimws(.))) %>%
  mutate(across(where(is.character), ~ na_if(., ""))) %>%
  
  # Drop records with NA values
  drop_na()
```

## Measurement level check

```{r}
# Convert variables to factor types
data <- data %>%
  dplyr::mutate(
    cancellation_policy = factor(cancellation_policy, 
                                 levels = c("flexible", "moderate", "strict"),
                                 ordered = TRUE),
    room_type = factor(room_type,
                       ordered = FALSE),
    review_rate_number = factor(review_rate_number,
                              levels = c(1, 2, 3, 4, 5),
                              ordered = TRUE)
  )

# View new structure
str(data$cancellation_policy)
str(data$room_type)
str(data$review_rate_number)
```

```{r}
# Parse price numeric values
data <- data %>%
  dplyr::mutate(
    price_usd = parse_number(price_usd),
    service_fee = parse_number(service_fee)
  )
```

## Individual outlier removal

```{r}
# Store data before outlier removal for later comparison
data_pre_outlier_removal <- data
```

### Functions

#### capitalize_str()

```{r}
# Function to format column name to spaced, capitalized string
capitalize_str <- function(string) {
  # Replace underscores with spaces and capitalize each word
  formatted_name <- gsub("_", " ", string)
  formatted_name <- tools::toTitleCase(formatted_name)
  
  # Return result
  return(formatted_name)
}
```

#### summary_table_cat()

```{r}
# Function to display comparison summary table
summary_table_cat <- function(data1, data2 = NULL, variable, sort = "none") {
  # Summarize data1
  summary_data1 <- data1 %>%
    group_by(across(all_of(variable))) %>%
    summarise(
      `Frequency (Before)` = n(),
      `Percentage (Before)` = (n() / nrow(data1)) * 100
    )
  
  if (!is.null(data2)) {
    # Summarize data2 if provided
    summary_data2 <- data2 %>%
      group_by(across(all_of(variable))) %>%
      summarise(
        `Frequency (After)` = n(),
        `Percentage (After)` = (n() / nrow(data2)) * 100
      )
    
    # Merge summaries
    result <- full_join(summary_data1, summary_data2, by = variable) %>%
      mutate(
        `Percentage (Before)` = sprintf("%.2f%%", `Percentage (Before)`),
        `Percentage (After)` = sprintf("%.2f%%", `Percentage (After)`)
      )
    
    # Apply sorting
    if (sort == "asc") {
      result <- result %>% arrange(`Frequency (Before)`)
    } else if (sort == "desc") {
      result <- result %>% arrange(desc(`Frequency (Before)`))
    }
  } else {
    # For single dataset case
    result <- summary_data1 %>%
      mutate(
        `Percentage (Before)` = sprintf("%.2f%%", `Percentage (Before)`)
      ) %>%
      rename(`Frequency` = `Frequency (Before)`,
             `Percentage` = `Percentage (Before)`)
    
    # Apply sorting
    if (sort == "asc") {
      result <- result %>% arrange(Frequency)
    } else if (sort == "desc") {
      result <- result %>% arrange(desc(Frequency))
    }
  }
  
  # Rename the variable column
  result <- result %>%
    rename(!!capitalize_str(variable) := all_of(variable))
  
  # Display table
  result <- result %>%
    kbl() %>%
    kable_classic() %>%
    kable_styling(latex_options = c("striped"), full_width = F)
  
  # Return table
  return(result)
}

```

#### summary_table_nums()

```{r}
# Create function to create summary table for numerical variables
summary_table_nums <- function(data, variables) {
  # Compile summary table
  summary <- data.frame(
    Mean = sapply(variables, function(var) round(mean(data[[var]]), 3)),
    Min = sapply(variables, function(var) min(data[[var]])),
    Median = sapply(variables, function(var) median(data[[var]])),
    Max = sapply(variables, function(var) max(data[[var]])),
    SD = sapply(variables, function(var) round(sd(data[[var]]), 3))
  )
  
  # Temporarily rename rownames using capitalize_str
  rownames(summary) <- sapply(variables, capitalize_str)
  
  # Display summary table
  summary <- summary %>%
    kbl() %>%
    kable_classic(latex_options = c("striped"), full_width = TRUE)
  
  # Return table
  return(summary)
}
```

#### plot_histogram()

```{r}
# Function to create a histogram
plot_histogram <- function(data, variable, title = NULL, binwidth = NULL, bins = NULL) {
  
  # Compile plot
  plot <- ggplot(data, aes(x = .data[[variable]])) +
    geom_histogram(binwidth = binwidth, 
                   bins = bins, 
                   fill = "lightblue",
                   color = "black") +
    labs(
      title = title,
      x = capitalize_str(variable),
      y = "Frequency",
      fill = variable
    ) +
    theme_bw()
  
  # Return plot
  return(plot)
}
```

### General

```{r}
summary_table_nums(data, c("construction_year", 
                               "service_fee", 
                               "minimum_nights", 
                               "number_of_reviews", 
                               "reviews_per_month", 
                               "availability_365", 
                               "price_usd")
                   )
```

### Categorical variables

#### 1: cancellation_policy

```{r}
# Verify structure of factors
str(data$cancellation_policy)

# Verify levels of factors
levels(data$cancellation_policy)

# Verify if ordered
is.ordered(data$cancellation_policy)
```

```{r}
# Show summary table
(table_1_pre <- summary_table_cat(data, 
                                  variable = "cancellation_policy")
)
```

```{r}
# Show bar plot
(plot_1_pre <- ggplot(data) +
  geom_bar(aes(x = cancellation_policy, fill = cancellation_policy)) +
  scale_fill_manual(values = c("flexible" = "lightgreen", 
                               "moderate" = "lightblue", 
                               "strict" = "tomato")) +
  labs(title = "Bar Plot of Cancellation Policy",
       fill = "Type of Policy",
       x = "Type of Policy",
       y = "Frequency"
      ) +
  theme_bw()
)
```

#### 2: room_type

```{r}
# Verify structure of factors
str(data$room_type)

# Verify levels of factors
levels(data$room_type)

# Verify if ordered
is.ordered(data$room_type)
```

```{r}
# Show summary table
(table_2_pre <- summary_table_cat(data, 
                                  variable = "room_type", 
                                  sort = "desc")
)
```

```{r}
# Show bar plot
# https://www.learnui.design/tools/data-color-picker.html#palette
(plot_2_pre <- ggplot(data) +
  geom_bar(aes(x = fct_infreq(room_type,         # Sort on freq
                              ordered = FALSE),
               fill = room_type)) +
  scale_fill_manual(values = c("Entire home/apt" = "#58508d", 
                               "Private room" = "#ff6361", 
                               "Shared room" = "#ffa600",
                               "Hotel room" = "#bc5090")) +
  labs(title = "Bar Plot of Room Type",
       fill = "Type of Room",
       x = "Type of Room",
       y = "Frequency"
      ) +
  theme_bw()
)
```

#### 3: review_rate_number

```{r}
# Verify structure of factors
str(data$review_rate_number_f)

# Verify levels of factors
levels(data$review_rate_number_f)

# Verify if ordered
is.ordered(data$review_rate_number_f)
```

```{r}
# Show summary table
(table_3_pre <- summary_table_cat(data, 
                                  variable = "review_rate_number")
)
```

```{r}
# Show bar plot
# https://www.learnui.design/tools/data-color-picker.html#divergent
(plot_3_pre <- ggplot(data) +
  geom_bar(aes(x = review_rate_number,
               fill = review_rate_number)) +
  scale_fill_manual(values = c("1" = "#de425b", 
                               "2" = "#fa9e5d", 
                               "3" = "#ffd582",
                               "4" = "#8bbe7a",
                               "5" = "#488f31")) +
  labs(title = "Before",
       fill = "Rate Number",
       x = "Rate Number",
       y = "Frequency"
      ) +
  theme_bw()
)
```

### Numerical variables

#### 4: construction_year

```{r}
# Plot histogram
(plot_4_pre <- plot_histogram(data, 
                              "construction_year", 
                              "Before",
                              binwidth = 1)
)
```

#### 5: service_fee

```{r}
# Plot histogram
(plot_5_pre <- plot_histogram(data, 
                              "service_fee", 
                              "Before",
                              binwidth = 5)
)
```

#### 6: minimum_nights

[ ] Removed values below 1 and above 365

```{r}
# Plot histogram
(plot_6_pre <- plot_histogram(data, 
                              "minimum_nights", 
                              "Before",
                              binwidth = 3)
 )
```

```{r}
# Count values
sum(data$minimum_nights < 1)
sum(data$minimum_nights > 365)
```

```{r}
# Remove all values below 1 and above 365 from dataset
data <- data[data$minimum_nights >= 1 & data$minimum_nights <= 365, ]
```

#### 7: number_of_reviews

```{r}
# Plot histogram
(plot_7_pre <- plot_histogram(data, 
                              "number_of_reviews", 
                              "Before",
                              binwidth = 10)
)
```

#### 8: reviews_per_month

[ ] Removed reviews_per_month > 45 (one and a half month)

```{r}
# Plot histogram
(plot_8_pre <- plot_histogram(data, 
                              "reviews_per_month", 
                              "Before",
                              binwidth = 1)
)
```

```{r}
# View count of reviews_per_month values
data %>%
  count(reviews_per_month) %>%
  arrange(desc(reviews_per_month))
```

```{r}
# Remove values > 45 reviews_per_month
data <- data[data$reviews_per_month <= 45, ]
```

#### 9: availability_365

```{r}
# Plot histogram
(plot_9_pre <- plot_histogram(data, 
                              "availability_365", 
                              "Before",
                              binwidth = 7)
)
```

```{r}
# Count values
sum(data$availability_365 < 0)
sum(data$availability_365 > 365)
```

```{r}
# Remove values below 1 and above 365
data <- data[data$availability_365 >= 0 & data$availability_365 <= 365, ]
```

#### 10: price_usd

```{r}
# Plot histogram
(plot_10_pre <- plot_histogram(data, 
                               "price_usd", 
                               "Before",
                               binwidth = 10)
)
```

## Final variable inspection

### Categorical variables

#### 1: cancellation_policy

```{r}
# View summary comparison of cancellation_policy
summary_table_cat(data1 = data_pre_outlier_removal, 
                  data2 = data,
                  variable = "cancellation_policy")
```


#### 2: room_type

```{r}
# View summary comparison of room_type
summary_table_cat(data1 = data_pre_outlier_removal, 
                  data2 = data,
                  variable = "room_type")
```


#### 3: review_rate_number

```{r}
# View summary comparison of review_rate_number
summary_table_cat(data1 = data_pre_outlier_removal, 
                  data2 = data,
                  variable = "review_rate_number")
```


### Numerical variables

```{r}
summary_table_nums(data, c("construction_year", 
                           "service_fee", 
                           "minimum_nights", 
                           "number_of_reviews", 
                           "reviews_per_month", 
                           "availability_365", 
                           "price_usd")
                   )
```

#### 4: construction_year

```{r}
# Plot histogram after outlier removal
plot_4_post <- plot_histogram(data, 
                              "construction_year", 
                              "After",
                              binwidth = 1)

# Compare plots of before and after outlier removal
ggarrange(plot_4_pre, plot_4_post, ncol = 2, nrow = 1)
```

#### 5: service_fee

```{r}
# Plot histogram after outlier removal
plot_5_post <- plot_histogram(data, 
                              "service_fee", 
                              "After",
                              binwidth = 5)

# Compare plots of before and after outlier removal
ggarrange(plot_5_pre, plot_5_post, ncol = 2, nrow = 1)
```

#### 6: minimum_nights

```{r}
# Plot histogram after outlier removal
plot_6_post <- plot_histogram(data, 
                              "minimum_nights", 
                              "After",
                              binwidth = 3)

# Compare plots of before and after outlier removal
ggarrange(plot_6_pre, plot_6_post, ncol = 2, nrow = 1)
```

#### 7: number_of_reviews

```{r}
# Plot histogram after outlier removal
plot_7_post <- plot_histogram(data, 
                              "number_of_reviews", 
                              "After",
                              binwidth = 10)

# Compare plots of before and after outlier removal
ggarrange(plot_7_pre, plot_7_post, ncol = 2, nrow = 1)
```

#### 8: reviews_per_month

```{r}
# Plot histogram after outlier removal
plot_8_post <- plot_histogram(data, 
                              "reviews_per_month", 
                              "After",
                              binwidth = 1)

# Compare plots of before and after outlier removal
ggarrange(plot_8_pre, plot_8_post, ncol = 2, nrow = 1)
```

#### 9: availability_365

```{r}
# Plot histogram after outlier removal
plot_9_post <- plot_histogram(data, 
                              "availability_365", 
                              "After",
                              binwidth = 7)

# Compare plots of before and after outlier removal
ggarrange(plot_9_pre, plot_9_post, ncol = 2, nrow = 1)
```

#### 10: price_usd

```{r}
# Plot histogram after outlier removal
plot_10_post <- plot_histogram(data, 
                              "price_usd", 
                              "After",
                              binwidth = 10)

# Compare plots of before and after outlier removal
ggarrange(plot_10_pre, plot_10_post, ncol = 2, nrow = 1)
```

# Linear Regression

## Model 1




