---
title: "FTDR Assignment 1"
author: "Jesse Nieuwkoop (1689959)"
date: "28-11-2024-"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

---

# Preparation

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
# Import the libraries
library(ggplot2)
library(tidyverse)
```

## Read data

```{r}
# Read data
data_raw <- read.csv("Airbnb_Open_Data.csv")

# View data structure
dplyr::glimpse(data_raw)
```

# Preprocessing

```{r}
# Load raw data onto new object
data <- data_raw
```

## Duplicates

```{r}
# Print total duplicates
sum(duplicated(data))

# Remove duplicate rows 
data <- data %>% unique()
```

## Subset & Rename

```{r}
# Subset data
data <- data %>%
  dplyr::select(cancellation_policy,
                room.type,
                Construction.year,
                service.fee,
                minimum.nights,
                number.of.reviews,
                reviews.per.month,
                review.rate.number,
                availability.365,
                price)

# Rename column names to snake_case format
data <- data %>%
  dplyr::rename(
    room_type = room.type,
    construction_year = Construction.year,
    service_fee = service.fee,
    minimum_nights = minimum.nights,
    number_of_reviews = number.of.reviews,
    reviews_per_month = reviews.per.month,
    review_rate_number = review.rate.number,
    availability_365 = availability.365,
    price_usd = price
  )
```

## Missing values

```{r}
# Count missing data
data %>%
  dplyr::summarise(across(everything(), ~ sum(is.na(.)))) %>%
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "NA Count"
  )
```

```{r}
# Drop all rows that contain empty or missing values
data <- data %>%
  # Convert empty character variables to NA
  mutate(across(where(is.character), ~ trimws(.))) %>%
  mutate(across(where(is.character), ~ na_if(., ""))) %>%
  
  # Drop records with NA values
  drop_na()
```

## Measurement level check and adjustment

```{r}
# Convert variables to factor types
data <- data %>%
  dplyr::mutate(
    cancellation_policy = factor(cancellation_policy, 
                                 levels = c("flexible", "moderate", "strict"),
                                 ordered = TRUE),
    room_type = factor(room_type,
                       ordered = FALSE)
  )

# View new structure
str(data$cancellation_policy)
str(data$room_type)
```

```{r}
# Parse price numeric values
data <- data %>%
  dplyr::mutate(
    price_usd = parse_number(price_usd),
    service_fee = parse_number(service_fee)
  )
```

## Individual variable check (Inspect)

Bedoeling om voor elke variabele een nuttige figuur weer te geven en summary descriptives te laten zien.

Categorisch:

- Counts + Percentages
- Plot: Bar plot

Continuous:

- Mean, median, range, min, max, sd
- Plot: Histogram
- Als eindig nummer (finite), dan mag ook bar plot.

Dingen die opvallen verwijderen.

GEEN IQR! Alleen werkelijk onmogelijk waarden weg.

### 1: cancellation_policy

```{r}
# Verify structure of factors
str(data$cancellation_policy)

# Verify levels of factors
levels(data$cancellation_policy)

# Verify if ordered
is.ordered(data$cancellation_policy)
```

### 2: room_type

### 3: review_rate_number

### 4: construction_year

### 5: service_fee

### 6: minimum_nights

### 7: number_of_reviews

### 8: reviews_per_month

### 9: availability_365

### 10: price_usd

# Linear Regression
