---
title: "assignment1_judith"
author: "j. van der Wolf"
date: "2024-11-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library
library(tidyverse)
library(dplyr)

getwd()
#data
data <- read.csv("../Airbnb_Open_Data.csv")
```

#Datacleaning 
## Duplicates
```{r}
# Print total duplicates
sum(duplicated(data$host.id))

# Remove duplicate rows 
data <- data %>% 
  dplyr::distinct(host.id, .keep_all = TRUE)
```

## Subset & Rename
```{r}
# Subset data
data <- data %>%
  dplyr::select(cancellation_policy,
                room.type,
                Construction.year,
                service.fee,
                minimum.nights,
                number.of.reviews,
                reviews.per.month,
                review.rate.number,
                availability.365,
                price)

# Rename column names to snake_case format
data <- data %>%
  dplyr::rename(
    room_type = room.type,
    construction_year = Construction.year,
    service_fee = service.fee,
    minimum_nights = minimum.nights,
    number_of_reviews = number.of.reviews,
    reviews_per_month = reviews.per.month,
    review_rate_number = review.rate.number,
    availability_365 = availability.365,
    price_usd = price
  )
```

## Missing values
```{r}
# Count missing data
data %>%
  dplyr::summarise(across(everything(), ~ sum(is.na(.)))) %>%
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "NA Count"
  )
```

```{r}
# Drop all rows that contain empty or missing values
data <- data %>%
  # Convert empty character variables to NA
  mutate(across(where(is.character), ~ trimws(.))) %>%
  mutate(across(where(is.character), ~ na_if(., ""))) %>%
  
  # Drop records with NA values
  drop_na()
```

## Measurement level check and adjustment
```{r}
# Convert variables to factor types
data <- data %>%
  dplyr::mutate(
    cancellation_policy = factor(cancellation_policy, 
                                 levels = c("flexible", "moderate", "strict")),
    room_type = factor(room_type),
   review_rate_number_f = factor(review_rate_number,
                                 labels = c("one", "two", "three", "four", "five")))
   
# View new structure
str(data$cancellation_policy)
str(data$room_type)
str(data$review_rate_number_f)
```

```{r}
# Parse price numeric values
data <- data %>%
  dplyr::mutate(
    price_usd = parse_number(price_usd),
    service_fee = parse_number(service_fee)
  )
```

#Individual variables 

##construction_year
```{r}
(summary_constr <- data %>%
  summarise(min(construction_year),
            max(construction_year),
            median(construction_year)))

#reken uit hoeveel observaties er zijn voor elk jaartal
data %>% 
  count(construction_year)

#percentages
table(data$construction_year)/ nrow(data)

#omdat er maar 20 jaartallen zijn; bar-chart laat visueel beter zien vind ik. 
ggplot(data, aes(construction_year)) +
  geom_bar()

```

##service_fee
```{r}
summary(data$service_fee)
ggplot(data, aes(service_fee)) +
  geom_histogram(binwidth = 0.3)
```

##minimum_nights
```{r}
summary(data$minimum_nights)
#Extreme low value (negative = not possibe) and extreme high value (5645 = ingevuld om andere reden dan om aan te geven wat minimale nachten zijn - bijv. voor een periode niet verhuren, maar wel omhoog blijven komen in zoeksuggesties e.d.)

#counting negative values (10)
data %>% 
  count(minimum_nights < 0)

#counting high values => bigger than 1 year, 365 = 17 (bigger than 1/2 year = 93)
data %>% 
  count(minimum_nights > 365)

data %>%
  filter(between(minimum_nights, 0, 365)) %>%
   summarise(min(minimum_nights),
            max(minimum_nights),
            mean(minimum_nights))

data %>%
  filter(between(minimum_nights, 0, 365)) %>%
  ggplot(aes(minimum_nights)) +
  geom_histogram(binwidth = 0.3)

```

```{r}
cor(data$price_usd, data$service_fee)
```

We were wondering what service_fee was (in relation to price). We checked the correlation between those two variables and found that they almost perfectly correlated. Therefore we made the decision to remove the variable service_fee from analysis to avoid issues related to multicollinearity. 


```{r}
data_cleaned <- data %>%
  select(-service_fee) %>%
   filter(between(minimum_nights, 0, 365)) %>%
  filter(between(availability_365, 0, 365)) %>%
  filter(price_usd > 0) 

str(data)

```

Check regression 
```{r}
model1 <- lm(price_usd ~ construction_year, data_cleaned)
summary(model1)

#center construction_year 
data_cleaned <- data_cleaned %>%
  mutate(construction_year_c = construction_year - 2003)

model2 <- lm(price_usd ~ construction_year_c, data_cleaned)
summary(model2)

(ggplot1 <- ggplot(data_cleaned, aes(construction_year_c, price_usd)) +
  geom_point()) +
  stat_smooth(method = "lm")
```

```{r}
model3 <- lm(price_usd ~ review_rate_number_f, data_cleaned)
summary(model3)

model4 <-lm(price_usd ~ room_type, data_cleaned)
summary(model4)

(ggplot2 <- ggplot(data_cleaned, aes(review_rate_number_f, price_usd)) +
  geom_point()) +
  stat_smooth(method = "lm")
```

