---
title: "FTDR Assignment 1: Logistic Regression"
author: "Kladbestand van Jesse Nieuwkoop (1689959)"
date: "02-12-2024"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      expand: 3
    df_print: paged
---

---

# Preparation

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
# Import the libraries
library(ggplot2)
library(tidyverse)

# 02/12/2024
library(kableExtra)
```

## Read data

```{r}
# Read data
data_raw <- read.csv("../Airbnb_Open_Data.csv")

# View data structure
dplyr::glimpse(data_raw)
```

# Preprocessing

```{r}
# Load raw data onto new object
data <- data_raw
```

## Duplicates

```{r}
# Print total duplicates
sum(duplicated(data))

# Remove duplicate rows 
data <- data %>% unique()
```

## Subset & Rename

```{r}
# Subset data
data <- data %>%
  dplyr::select(cancellation_policy,
                room.type,
                Construction.year,
                service.fee,
                minimum.nights,
                number.of.reviews,
                reviews.per.month,
                review.rate.number,
                availability.365,
                price)

# Rename column names to snake_case format
data <- data %>%
  dplyr::rename(
    room_type = room.type,
    construction_year = Construction.year,
    service_fee = service.fee,
    minimum_nights = minimum.nights,
    number_of_reviews = number.of.reviews,
    reviews_per_month = reviews.per.month,
    review_rate_number = review.rate.number,
    availability_365 = availability.365,
    price_usd = price
  )
```

## Missing values

```{r}
# Count missing data
data %>%
  dplyr::summarise(across(everything(), ~ sum(is.na(.)))) %>%
  tidyr::pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "NA Count"
  )
```

```{r}
# Drop all rows that contain empty or missing values
data <- data %>%
  # Convert empty character variables to NA
  mutate(across(where(is.character), ~ trimws(.))) %>%
  mutate(across(where(is.character), ~ na_if(., ""))) %>%
  
  # Drop records with NA values
  drop_na()
```

## Measurement level check and adjustment

```{r}
# Convert variables to factor types
data <- data %>%
  dplyr::mutate(
    cancellation_policy = factor(cancellation_policy, 
                                 levels = c("flexible", "moderate", "strict"),
                                 ordered = TRUE),
    room_type = factor(room_type,
                       ordered = FALSE),
    review_rate_number_f = factor(review_rate_number,
                                  levels = c(1, 2, 3, 4, 5),
                                  ordered = TRUE)
  )

# View new structure
str(data$cancellation_policy)
str(data$room_type)
```

```{r}
# Parse price numeric values
data <- data %>%
  dplyr::mutate(
    price_usd = parse_number(price_usd),
    service_fee = parse_number(service_fee)
  )
```

## Individual variable check (Inspect)

Bedoeling om voor elke variabele een nuttige figuur weer te geven en summary descriptives te laten zien.

Categorisch:

- Counts + Percentages
- Plot: Bar plot

Continuous:

- Mean, median, range, min, max, sd
- Plot: Histogram
- Als eindig nummer (finite), dan mag ook bar plot.

Dingen die opvallen verwijderen.

GEEN IQR! Alleen werkelijk onmogelijk waarden weg.

### 1: cancellation_policy

```{r}
# Verify structure of factors
str(data$cancellation_policy)

# Verify levels of factors
levels(data$cancellation_policy)

# Verify if ordered
is.ordered(data$cancellation_policy)
```

```{r}
# Function to display summary table
summary_table_cat <- function(data, variable, name, sort = "none") {
  result <- data %>%
    group_by(across(all_of(variable))) %>%
    summarise(Count = n(),
      Percentage = (n() / nrow(data)) * 100
    ) %>%
    mutate(Percentage = sprintf("%.2f%%", Percentage)) %>%
    rename(!!name := .data[[variable]])
  
  # Apply sorting
  if (sort == "asc") {
    result <- result %>% arrange(Count)
  } else if (sort == "desc") {
    result <- result %>% arrange(desc(Count))
  }
  
  # Print result
  result %>%
    kbl() %>%
    kable_classic() %>%
    kable_styling(latex_options = c("striped"), full_width = F)
}
```

```{r}
# Show summary table
summary_table_cat(data, "cancellation_policy", "Type of Cancellation Policy")
```

```{r}
# Show bar plot
ggplot(data) +
  geom_bar(aes(x = cancellation_policy, fill = cancellation_policy)) +
  scale_fill_manual(values = c("flexible" = "lightgreen", 
                               "moderate" = "lightblue", 
                               "strict" = "tomato")) +
  labs(title = "Bar Plot of Cancellation Policy",
       fill = "Type of Policy",
       x = "Type of Policy",
       y = "Count"
      ) +
  theme_bw()
```

### 2: room_type

```{r}
# Verify structure of factors
str(data$room_type)

# Verify levels of factors
levels(data$room_type)

# Verify if ordered
is.ordered(data$room_type)
```
```{r}
# Show summary table
summary_table_cat(data, "room_type", "Type of Room", "desc")
```

```{r}
# Show bar plot
# https://www.learnui.design/tools/data-color-picker.html#palette
ggplot(data) +
  geom_bar(aes(x = fct_infreq(room_type,         # Sort on freq
                              ordered = FALSE),
               fill = room_type)) +
  scale_fill_manual(values = c("Entire home/apt" = "#58508d", 
                               "Private room" = "#ff6361", 
                               "Shared room" = "#ffa600",
                               "Hotel room" = "#bc5090")) +
  labs(title = "Bar Plot of Room Type",
       fill = "Type of Room",
       x = "Type of Room",
       y = "Count"
      ) +
  theme_bw()
```

### 3: review_rate_number

#### Continuous: review_rate_number

```{r}
# Function to summarize numerical variables
summary_table_num <- function(data, variable, name) {
  summary <- data %>%
    summarise(
      Mean = round(mean(.data[[variable]], na.rm = TRUE), 3),
      Min = min(.data[[variable]], na.rm = TRUE),
      Median = median(.data[[variable]], na.rm = TRUE),
      Max = max(.data[[variable]], na.rm = TRUE),
      SD = round(sd(.data[[variable]], na.rm = TRUE), 3)
    ) %>%
    mutate(Variable = name, .before = everything())  # Add Variable column 
  
  # Print the summary table
  summary %>%
    kbl() %>%
    kable_classic() %>%
    kable_styling(latex_options = c("striped"), full_width = F)
}
```

```{r}
# Show descriptive summary
summary_table_num(data, "review_rate_number", "Review Rate Number")
```

```{r}
# Plot histogram
ggplot(data, aes(x = review_rate_number)) +
  geom_histogram(binwidth = 1, 
                 fill = "lightblue",
                 color = "black") +
  labs(
    title = "Histogram of Review Rate Number",
    x = "Rate Number",
    y = "Count",
    fill = "Rate Number"
  ) +
  theme_bw()

```

#### Ordinal: review_rate_number_f

```{r}
# Verify structure of factors
str(data$review_rate_number_f)

# Verify levels of factors
levels(data$review_rate_number_f)

# Verify if ordered
is.ordered(data$review_rate_number_f)
```

```{r}
# Show summary table
summary_table_cat(data, "review_rate_number_f", "Rate Number")
```

```{r}
# Show bar plot
# https://www.learnui.design/tools/data-color-picker.html#divergent
ggplot(data) +
  geom_bar(aes(x = review_rate_number_f,
               fill = review_rate_number_f)) +
  scale_fill_manual(values = c("1" = "#de425b", 
                               "2" = "#fa9e5d", 
                               "3" = "#ffd582",
                               "4" = "#8bbe7a",
                               "5" = "#488f31")) +
  labs(title = "Bar Plot of Review Rate Number",
       fill = "Rate Number",
       x = "Rate Number",
       y = "Count"
      ) +
  theme_bw()
```

### 4: construction_year

### 5: service_fee

### 6: minimum_nights

### 7: number_of_reviews

### 8: reviews_per_month

### 9: availability_365

### 10: price_usd

# Linear Regression
